# Summary

```{r, echo=F, warning=F, message=F, include=F}
source("R-code/energy-analysis.R")
```

## Introduction {.unnumbered}

With climate change on the rise, the Dauphin Island Power Plant must consider how increasing water temperatures will impact its operations. Currently, the plant relies on the assumption that water temperatures will remain below 85째F. When temperatures exceed this threshold, the power plant must operate at reduced capacity, limiting its ability to meet energy demands. **Predicting future water temperatures and calculating the likelihood that they exceed this threshold is critical to mitigating risks.** Furthermore, examining how periods of high water temperature correspond to peaks in energy demand will help us better evaluate the impact of reduced operating capacity.

## Results {.unnumbered}

Using a standard linear model, water temperatures were projected for the next five years (2025-2029) on the coast of Dauphin Island. **Eighteen months over the next five years are predicted to have water temperatures exceeding the threshold,** almost all with a probability approaching 100%. The months that are projected to repeatedly exceed 85째F are **June, July, August, and September.**

In the secondary analysis of energy demand, monthly averages were examined to find when energy demand was at its peak. **Alarmingly, there was a huge spike in energy demand in the months that we identified would exceed the water temperature threshold.** Specifically, July and August have the highest energy demand out of the whole year. Monthly energy demand is presented in Figure 1, where the blue shaded region indicates months when projected water temperatures are expected to exceed 85째F.

```{r, echo=F}
ggplot(monthly_data, aes(x=as.numeric(Month), y=Pred_Demand)) +
  geom_line() +
  annotate("rect",
           xmin = 6, xmax = 9,
           ymin = -Inf, ymax = +Inf,
           fill = "steelblue", alpha = 0.2 ) +
  scale_x_continuous(breaks=1:12, labels=month.abb) +
  labs(title = "Energy Demand By Month",
       x = "Month", y = "Energy Demand (MWh)")

```

| **Figure 1:** *Energy Demand By Month.* Month is on the x-axis and Energy Demand (megawatt-hours) is on the y-axis. The blue shaded region corresponds to the months where water temperatures are predicted to exceed the 85째F threshold.

## Discussion {.unnumbered}

These findings raise substantial concern for the Dauphin Island Power Plant. One-third of each upcoming year is almost guaranteed to have water temperatures exceeding the threshold. As a result, **the plant will be operating at reduced capacity for a third of each year in the near future.** Compounding this issue, periods of high water temperatures align with times of high energy demand, magnifying the impact of reduced capacity.

While these results are troubling, they are not surprising as climate change has greatly impacted sustainability across most industries. The hottest months of the year were identified in our analysis, which only makes sense that they correspond to the highest water temperatures. **The Dauphin Island Power Plant needs to immediately explore different solutions to combat these rising temperatures, such as raising the temperature threshold at which reduced operating capacity kicks in. Starting preparation now is essential to maintaining reliable energy production in the coming years.**
